---
title: "p8105_hw2_wb2366"
auther: Wenhan Bao
Date: 2021-10-09
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, results='hide'}
library(tidyverse)
library(readxl)
library(lubridate)
library(dplyr)
library(ggplot2)
```

## Problem1

**Read and clean the Mr. Trash Wheel sheet**

```{r import and clean data}
Trash_wheel_data = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                   sheet = "Mr. Trash Wheel",
                   range = "A2:N534") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls))
```

**Read and clean precipitation data for 2018 and 2019.**

```{r read and clean preciptation}
# Import 2019 Precipitation data
precipitation_2019 = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                   sheet = "2019 Precipitation",
                   range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = "2019")

# Import 2018 Precipitation data
precipitation_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                   sheet = "2018 Precipitation",
                   range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = "2018")

# Combine two datasets
prec_2018_2019 = bind_rows(precipitation_2018, precipitation_2019) %>%
  janitor::clean_names() %>% 
  mutate(month = month.name[month]) %>% 
  relocate("year", "month", "total")
```

**Write a paragraph about these data**

**Answer:** For the first data set, Mr. Trash Wheel contains ``r nrow(Trash_wheel_data)`` observations with ``r ncol(Trash_wheel_data)`` variables which are respectively `Trash_wheel_data`(``r names(Trash_wheel_data)``)

For the second data set, precipitation data for 2018 and 2019 contains ``r nrow(prec_2018_2019)`` observations with ``r ncol(prec_2018_2019)`` variables which are respectively `prec_2018_2019`(``r names(prec_2018_2019)``)

**what was the total precipitation in 2018?**

**Answer:** The total precipitation in 2018 is `r filter(prec_2018_2019, year == "2018") %>% pull(total) %>% sum()`

**What was the median number of sports balls in a dumpster in 2019?**

**Answer:** The median number of sports balls in a dumpster in 2019 is `r filter(Trash_wheel_data, year == "2019") %>% pull(sports_balls) %>% median()`

## Problem2

**clean the data in pols-month.csv.**
```{r clean_pols_month}
pols_month = read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv")%>%
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(year = as.character(year),
         month = as.integer(month),
         month = month.name[month],
         president = ifelse(prez_gop == 0, "dem", "gop")) %>% 
  select(-"prez_dem" & -"prez_gop" & -"day") %>% 
  relocate("year", "month", "president")
```

**clean the data in snp.csv**
```{r clean_snp_data}
snp_data = read_csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.character(year),
         month = as.integer(month),
         month = month.name[month]) %>% 
  select(-"day") %>% 
  relocate("year", "month")
```

**tidy the unemployment data**
```{r tidy_unemployment_data}
unemployment_data = read_csv(file = "./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec, 
               names_to = "month", 
               values_to = "unemployment") %>% 
  janitor::clean_names() %>% 
  mutate(month = match(month, month.abb),
         month = month.name[month],
         year = as.character(year))
```

**Join the datasets by merging snp into pols, and merging unemployment into the result.**
```{r join the datasets}
Pols_snp_unemploy_data <- 
  left_join(pols_month, snp_data, by = c("year", "month")) %>% 
  left_join(unemployment_data, by = c("year", "month"))
```

**Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset** 

**Answer:** For the pols-months data set, it contains ``r nrow(pols_month)`` observations with ``r ncol(pols_month)`` variables which are respectively ``r names(pols_month)``. The range of years is from 1947 to 2015.

For the snp data set, it contains ``r nrow(snp_data)`` observations with ``r ncol(snp_data)`` variables which are respectively ``r names(snp_data)``. The range of years is from 1969 to 2015.

For the unemployment data set, it contains ``r nrow(unemployment_data)`` observations with ``r ncol(unemployment_data)`` variables which are respectively ``r names(unemployment_data)``. The range of years is from 1948 to 2015.

For the merging data set, it contains ``r nrow(Pols_snp_unemploy_data)`` observations with ``r ncol(Pols_snp_unemploy_data)`` variables which are respectively ``r names(Pols_snp_unemploy_data)``

## Problem3

**Load and tidy the data.**
```{r load and tidy}
babynames_df = read_csv(file = "./data/Popular_Baby_Names.csv")%>%
  janitor::clean_names() %>% 
  mutate(
    childs_first_name = toupper(childs_first_name),
         ethnicity = gsub("BLACK NON HISP$", "BLACK NON HISPANIC", pull(.,ethnicity)),
         ethnicity = gsub("WHITE NON HISP$", "WHITE NON HISPANIC", pull(.,ethnicity)),
         ethnicity = gsub("ASIAN AND PACI$", "ASIAN AND PACIFIC ISLANDER", pull(.,ethnicity))) %>% 
  distinct()
```

**Produce a well-structured, reader-friendly table**
```{r Olivia_df}
Olivia_df = filter(babynames_df, childs_first_name == "OLIVIA" & gender == "FEMALE") %>% 
  select(-"count") %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>% 
  relocate(.after = c(1:3), (9:4))
  
  options(knitr.kable.NA = "")
  knitr::kable(Olivia_df, 
               caption = "The rank of “Olivia” as a female baby name over time")
```


**Produce a similar table showing the most popular name among male children over time.**
```{r popular name among male children}
popname_male_df = filter(babynames_df, gender == "MALE" & rank == "1") %>% 
  select(-"count" & -"rank") %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  ) %>% 
  relocate(.after = c(1:2), (8:3))
  
  options(knitr.kable.NA = "")
  knitr::kable(popname_male_df, 
               caption = "The most popular name among male children over time")
```

**Produce a scatter plot**
```{r scatter plot}
scatterplot_df = filter(babynames_df,
                        gender == "MALE" &
                        ethnicity == "WHITE NON HISPANIC" &
                        year_of_birth == "2016")

ggplot(scatterplot_df, aes(x = rank, y = count)) + 
  geom_point() +
  labs(title = "The number and rank of name for male, white non-hispanic children born in 2016", 
       x = "The rank in popularity of that name", 
       y = "The number of children with a name")
```






